Measurement Bias from Address Aliasing

### 2016-05-15 presentation.tex
Almost something to keep

\begin{itemize}
  \item Identify address aliasing as underlying explanation for measurement bias.
  % \begin{itemize}
  %   \item Looking at example from previous work
  % \end{itemize}
  \item See how aliasing affects dynamic memory allocation.
  \item Identify opportunities to optimize for aliasing.
\end{itemize}

@@

\pgfplotstableread{bin/microkernel-cycles-haswell.dat}{\stackoffsettable}
\begin{figure}
  \begin{tikzpicture}
    \begin{axis}[
        width = \textwidth,
        height = 5.0cm,
        font = \footnotesize,
        xlabel=Bytes added to environment,
        ylabel=Cycles,
        domain = 0:8192,
        xtick = {0,1024,...,8192},
        xmin = 0,
        xmax = 8192,
        cycle list name = exotic
      ]
      \addplot[ycomb] table[x expr = \thisrowno{0}*16, y = cycles:u] \stackoffsettable ;
    \end{axis}
  \end{tikzpicture}
  \caption{Measured average of 10 cycle count samples for 512 different environments; offseting stack by multiples of 16 bytes.}
\end{figure}

@@

% More generic on methodology? Not necessary for presentation maybe..
% Just to straight to the juice.
\begin{frame}
  \frametitle{How to Measure Address Aliasing}
  \framesubtitle{Using Hardware Performance Counters}

  % Some points on hardware perf counters.
  \begin{itemize}
    \item Control execution environment (disable ASLR).
    \item Using processor-specific instrumentation, we can \emph{count} various low-level events during execution.
      \begin{itemize}
        \item Cycles executed, branch prediction misses, instructions fetched, etc.
      \end{itemize}
    %\item Intel manual lists hundreds of events.
  \end{itemize}

  Events covering 4K aliasing:

  \begin{description}
    \item[{\small LD\_BLOCKS\_PARTIAL.ADDRESS\_ALIAS.}] 
    \emph{``Counts the number of loads that have partial address match with preceding stores, causing the load to be reissued.''} 
    \cite[B.3.4.4]{OptimizationManual}
  \end{description}
  %This event is listed in the manual for microarchitectures going back to ``Nehalem'', including ``Ivy Bridge'' and ``Haswell'' CPUs~\cite{Volume3B}.
  %Older architectures such as Core do not have this particular event listed, but 4K aliasing is covered by a more general event:
  \begin{description}
    \item[{\small LOAD\_BLOCKS.OVERLAP\_STORE.}] 
    \emph{``Loads that partially overlap an earlier store, or 4-Kbyte aliased with a previous store.''} 
    \cite[Table 19-17]{Volume3B}
  \end{description}

@@

Since page size is usually 4096 byte, such allocations will \emph{always} alias with each other.

@@

\begin{frame}
\frametitle{Optimizing for address alias}

@@

\begin{itemize}
  \item Mark buffers with \texttt{restrict}.
  \item Use a special purpose allocator.
  \item Manually adjust address offset.
\end{itemize}

>>>>

\begin{itemize}
  \item Identify address aliasing as an underlying explanation for measurement bias.
  \item See how aliasing affects dynamic memory allocation.
  \item Look at opportunities to optimize for aliasing.
\end{itemize}

@@

\pgfplotstableread{bin/microkernel-cycles-haswell.dat}{\stackoffsettable}
\begin{figure}
  \begin{tikzpicture}
    \begin{axis}[
        title = Haswell,
        width = \textwidth,
        height = 4.7cm,
        font = \footnotesize,
        xlabel=Bytes added to environment,
        ylabel=Cycles,
        domain = 0:8192,
        xtick = {0,1024,...,8192},
        xmin = 0,
        xmax = 8192,
        cycle list name = exotic
      ]
      \addplot[ycomb] table[x expr = \thisrowno{0}*16, y = cycles:u] \stackoffsettable ;
    \end{axis}
  \end{tikzpicture}
  \caption{Measured cycle count samples for 512 different environments.}
\end{figure}

@@

\begin{frame}[fragile]
  \frametitle{How to Measure Address Aliasing}

  % Some points on hardware perf counters.
  \begin{itemize}
    \item Control execution environment (disable ASLR).
    \item Sample alias occurrence using \emph{hardware performance counters}.
      % \begin{itemize}
      %   \item Cycles executed, branch prediction misses, instructions fetched, etc.
      % \end{itemize}
    %\item Intel manual lists hundreds of events.
  \end{itemize}

  \begin{description}
    \item[{\small LD\_BLOCKS\_PARTIAL.ADDRESS\_ALIAS.}] 
    \emph{``Counts the number of loads that have partial address match with preceding stores, causing the load to be reissued.''} 
    \cite[B.3.4.4]{OptimizationManual}
    Listed in the manual for microarchitectures going back to ``Nehalem'', including ``Ivy Bridge'' and ``Haswell'' CPUs~\cite{Volume3B}
  \end{description}
  %This event is listed in the manual for microarchitectures going back to ``Nehalem'', including ``Ivy Bridge'' and ``Haswell'' CPUs~\cite{Volume3B}.
  %Older architectures such as Core do not have this particular event listed, but 4K aliasing is covered by a more general event:
  \begin{description}
    \item[{\small LOAD\_BLOCKS.OVERLAP\_STORE.}] 
    \emph{``Loads that partially overlap an earlier store, or 4-Kbyte aliased with a previous store.''} 
    \cite[Table 19-17]{Volume3B}
    Covers 4K aliasing for older architectures, such as Core.
  \end{description}

@@

Since page size is 4 KiB, such allocations will \emph{always} alias with each other.

@@

\begin{frame}
\frametitle{Optimizing for address aliasing}

@@

\begin{itemize}
  \item Mark buffers with \texttt{restrict}.
  \begin{itemize}
    \item Can allow the optimizer to reduce number of memory accesses.
  \end{itemize}
  \item Use a special purpose allocator.
  \begin{itemize}
    \item Heuristics for avoiding pairwise aliasing buffers.
    \item No known implementation with this goal in mind, to our knowledge.
  \end{itemize}
  \item Manually adjust address offset.
  \begin{itemize}
    \item Consider memory address alignment for performance tuning inner loops.
  \end{itemize}
\end{itemize}
